We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-03-01 paper/paper.tex
edit

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical Dark Matter (DM) halo mass in the LG samples.
However, there are three common trends.
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in the M31 galaxy have a rather typical
distribution in the LCDM context, while the MW distribution is
atypical in this respect.  
These quantitative results place the LG at the level of a $3\sigma$
outlier in the LCDM paradigm. 
We suggest that understanding the reasons for this atipicality
will require quantifying the asphericity probability distribution as a
function of halo mass and large scale environment.
The approach presented here can facilitate that kind study and other
comparisons between different numerical setups and choices to study
satellites around LG pairs in simulations.  
\end{abstract}

@@

In the case of M31, observational studies have found a planar satellite
distribution in a special subset of $15$ satellites out of the 
total poulation of $27$ satellites \citep{2013ApJ...766..120C,
2013Natur.493...62I}, although considering the satellites ranked by
luminosity does not show any special feature.
All published studies agree on the point that the spatial distribution of
the 15 to 27 brightest satellites are consistent with  spherically
symmetric distribution and easy to reproduce in LCDM simulations
\citep{2006AJ....131.1405K,2007MNRAS.374.1125M, 2013ApJ...766..120C}.    
Table \ref{table:M31} and Table \ref{table:MW} summarize some more
details on the results we have just mentioned \footnote{
Those tables also include the results from this paper using the
methodolgy we describe in the upcoming sections.}.

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2006AJ....131.1405K}} & RMS height in 15 brightest
satellites & Monte Carlo satellite distributions with a power law
radial distribution & $87-99$\\
\text{\cite{2007MNRAS.374.1125M}} & Triaxiality and RMS height in 16 brightest satellites & Monte Carlo from an spherical power law radial
distribution & $17$\\
\text{\cite{2013ApJ...766..120C}}& RMS height in the 27 brightest
satellites & Monte Carlo randomized satellite distribution & High\\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body Dark Matter only simulation ($\sim
10^{6}$\Msun particle mass resolution)& 56 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 20 halos from selected pairs in a cosmological N-body hydro simulation ($\sim
10^{6}$\Msun particle mass resolution)& 27 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 37 \\
\hline
\end{tabular}
\caption{Probability to find  the triaxiality and/or root mean squared
  (RMS) height of M31 satellites in LCDM simulations. 
\label{table:M31}}
\end{table*}

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2005A&A...431..517K}} & RMS height in 11 classical
satellites & Monte Carlo from an spherical power law
radial distribution. & $<0.5$ \\
\text{\cite{2005MNRAS.363..146L}} & Triaxiality of 11 classical
satellites & 6 high resolution DM only N-body simulations ($\sim10^5$
\Msun\ particle mass resolution). & High\\ 
\text{\cite{2005ApJ...629..219Z}} & Triaxiality and disk height in 
11 classical satellites & 3 high resolution DM only N-body
simulations. ($\sim10^6$ \Msun\ particle mass resolution) & 2 \\
\text{\cite{2007MNRAS.374.1125M}} & Triaxiality and RMS height in 11-13 brightest satellites & Monte Carlo from a halo triaxiality distribution from LCDM
simulations & $<0.5$\\
\text{\cite{2009MNRAS.399..550L}}& Triaxiality in 11 classical satellites & 436 halos from a
cosmological N-body simulation ($1.3\times 10^{8}$\Msun\ particle mass)
&  High \\
\text{\cite{2013MNRAS.429..725S}}& $c/a$ ratio in 12 brightest
satellites & 6 High resolution DM only N-body simulations ($\sim
10^3$\Msun particle mass resolution) of individual halos & Low \\
\text{\cite{2013MNRAS.429.1502W}}& $c/a$ ratio or RMS height in 11 brightest
satellites & 1686 halos from a cosmological DM only N-body simulation
($\sim 10^6$\Msun particle mass resolution) & 6 or 13 \\
\text{\cite{2014ApJ...789L..24P}}& $c/a$ and $b/a$ ratio in 11
classical satellites & 48 high resoltion DM only N-body simulations
($\sim 10^{5}$\Msun particle mass resolution) of both halo pairs and
isolated halos & 0.77\\
\text{\cite{2016MNRAS.457.1931S}}& $c/a$ ratio in 11 classical satellites & 12
high resolution Hydro simulation of halo pairs ($\sim 10^{4}$\Msun particle
mass resolution) & High\\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body
Dark Matter only simulation ($\sim 10^{6}$\Msun particle mass resolution)& 3\\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 20 halos from selected pairs in acosmological N-body
hydro simulation ($\sim 10^{6}$\Msun particle mass resolution)& 0.6 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 0.2 \\
\hline
\end{tabular}
\caption{Same as Table \ref{table:M31} for the MW satellites.
\label{table:MW}}
\end{table*}

@@

We control the first effect by setting as a direct point of reference
an spherical satellite distribution and not the simulations themselves.
The spherical satellite distribution is built from the data itself
(observational or simulated) by randomizing the angular position
of each satellite around the central galaxy and keeping its radial
distance fixed \citep{2017AN....338..854P}. 
We characterize the satellites in terms of the scalars describing its
deviation from the spherical distribution. 
We then build an explicit analytic probability distribution for the
asphericity; this solves the problem of having a common reference
point to compare simulations.  
We use a simulation to estimate the parameters in this distribution to
later use it as a parent sample to generate any desired number of
samples that are by construction statistically compatible with the
simulation; thus overcoming the problem of having a small number of
systems in the parent simulations.

@@

The rest of the paper describes in detail our implementation and
results. It is structured as follows. 
In Section \ref{sec:DataSamples} we list the sources of the observational and
simulated data to be used throughout the paper.
In Section \ref{sec:SpatialMeasurements} we describe the methods we
use to build halo pairs and  
characterize its satellite distributions.
In Section \ref{sec:results} we present our results to finally
conclude in Section \ref{sec:conclusions}.

@@

The base for our analysis is the catalog compiled by
\cite{2014yCat..74351928P} which reports information on all 
galaxies within 3 Mpc around the Sun to that date. 
Detailed description of the compiled catalog can be found in
\cite{2013MNRAS.435.1928P}, here we summarize the relevant features
for the current study.
The information in the catalogue is based on the catalogue compiled by
\cite{2012AJ....144....4M}.
The distance estimates are based on resolved stellar populations. 
We use three dimensional positions in a cartesian coordinate system
as computed by \cite{2013MNRAS.435.1928P}.
In this coordinate system the $z$-axis points towards the Galactic north pole, the
$x$-axis points in the direction from the Sun to the Galactic centre,
and the $y$-axis points in the direction of the Galactic rotation.

For both the M31 and MW we only use the 11 to 15 brightest satellites (using
$M_V$ magnitudes) within a distance of $300$kpc to its central galaxy.
The satellites included for the MW analysis are: 
LMC, SMC, Canis Major, Sagittarius dSph, Fornax, Leo I, Sculptor,
Leo II, Sextans I, Carina, Ursa Minor, Draco, Canes Venatici (I),
Hercules and Bootes II.
The satellites included for the M31 analysis are: Triangulum, NGC205,
M32, IC10, NGC185, NGC147, Andromeda VII, Andromeda II, Andromeda
XXXII, Andromeda XXXI, Andromeda I, Andromeda VI, Andromeda XXIII, LGS 3, 
 and Andromeda III.

@@

We use publicly available data from the Illustris Project 
\citep{2014MNRAS.444.1518V}. 
This suite of cosmological simulations, performed using the quasi-Lagrangian
code AREPO \citep{2010MNRAS.401..791S}, followed the coupled evolution of dark 
matter and gas and includes parametrizations to account for the effects of
gas cooling, photoionization, star formation, stellar feedback, black
hole and super massive black hole feedback. 
The simulation volume is a cubic box of $75$ \Mpch\ on a side.
The cosmological parameters correspond to a $\Lambda$CDM cosmology
consistent with WMAP-9 measurements \citep{2013ApJS..208...19H}.

@@

We buid a sample of pairs that resemble the conditions in the LG.
To construct this sample we select from Illustris-1 all galaxies with a stellar
mass in the range $1\times10^{10}\Msun <M_{\star}<1.5 \times 10^{11}
\Msun$.
Then we select the pairs with the following conditions.

@@

In Illustris-1-Dark we use the center of mass position of the 27 pairs
in Illustris-1 to find the matching halo pairs.
After discarding the pairs with less than 15 detected subhalos in one
of the halos we end up with a total of 24 pairs in Illustris-1-Dark. 
This corresponds to a pair number density of $\sim 2 \times10^{-5}$
pairs Mpc$^{-3}$ 
Appendix A shows the physical  properties (stellar masses, maximum
circular velocities, radial velocities and separation) of those
pairs.

@@

We compute the satellite statistics for 11 up to 15 satellites.
The lower limit corresponds to the number of classical MW satellites,
while the upper limit corresponds to the minimum number of M31
satellites included in the analysis.
In simulations we rank the subhalos by their maximum circular
velocity, in observations we rank the satellites by its $M_V$
magnitude.

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Observations}}
    & \multicolumn{2}{c|}{\textbf{Randomized Obs.}} &
    \multicolumn{2}{c|}{\textbf{Normalized Units}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} \\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $59\pm 3$  & $21\pm 2$  & $65\pm 12$   &
    $45\pm 8$    & $-0.48\pm 0.24$ & $-2.48\pm 0.26$\\\hline
    $c/a$ ratio & $0.45\pm 0.04$ & $0.25\pm 0.05$ & $0.55\pm0.10$ &
    $0.53\pm 0.10$ & $-1.03\pm 0.37$ & $-2.18\pm 0.42$\\ \hline
    $b/a$ ratio & $0.82\pm 0.06$ & $0.80\pm 0.04$ & $0.82\pm0.07$ &
    $0.81\pm 0.08$ & $-0.02\pm 0.82$ & $-0.13\pm 0.47$\\ \hline
  \end{tabular}
  \caption{Results from observations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first colum refers to the results from
    observational data, the second column uses the spherically
    randomized version of the observational data, the third column
    corresponds to the observational data recentered and normalized
    by the randomized values. \label{table:observations}}
\end{table*}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Illustris-1-Dark}} & \multicolumn{2}{c|}{\textbf{Illustris-1}} & \multicolumn{2}{c|}{\textbf{ELVIS}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW}& \textbf{M31} & \textbf{MW}\\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $62\pm 3$   & $61\pm 3$     & $70\pm 4$ & $67\pm 2$ & $70\pm 2$& $68\pm 4$ \\\hline
    $c/a$ ratio & $0.50\pm0.01$ & $0.50\pm 0.02$ & $0.52\pm 0.01$ & $0.53\pm 0.01$ & $0.54\pm 0.01$& $0.49\pm 0.02$ \\ \hline
    $b/a$ ratio & $0.79\pm0.01$ & $0.81\pm 0.02$ & $0.80\pm 0.01$ & $0.80\pm 0.02$ & $0.80\pm0.01$& $0.81\pm 0.01$\\ \hline
  \end{tabular}
  \caption{Results from simulations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first column summarize the results from the 24 pairs in
    Illustris-1-Dark, the second column from the 20 pairs in
    Illustris-1 and the third column from the 12 pairs in the
    ELVIS project.\label{table:simulations}}
\end{table*}

@@

We compare each satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
For each quantity we compute the average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.
These normalized quantities are used to build the explicit probability
distributions for the scalars describing aesphericity.

@@

After building the normalized variables with the simulated data we
perform a Kolmogorov-Smirnov test with the null hypothesis of belonging
to a normal distribution with mean and standard deviation computed
from the mean and standard deviation estimated from the data.
Although the physical quantities of interests are bound, unlike the
random variables from a normal distribution, we find that the
distributions for the normalized $w$, $c/a$ and $b/a$ are indeed
consistent with gaussian ditributions.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in observations and
  simulations (Illustris-1-Dark in this case). In all panels the
  horizontal/vertical axis corresponds to the  M31/MW or the
  most/less massive halo in the pair 
  Left. Plane width in physical units comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions (circles). 
Middle. Average from the observations (star) and the average from each
pair in the simulation (circles).
Right. Same as the middle panel except that this time each point has
been normalized (median substracted and normalized by the standard
deviation) to the results of its randomization. 
The main message of these panels is that the MW has a significantly
thinner plane both compared to the result of its own satellite
spherical randomization (left panel) and the expectation from
simulations (middle and right panel).  
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In M31 the satellite distribution is in agreement with the
expectations both from an spherical distribution and the results form the
simulations. 
\label{fig:scatter_width}}
\end{figure*}

@@

Figure \ref{fig:correlations_illustrisdm} illustrates the results from
 computing the covariance matrix and mean vector in
 Eq.\ref{eq:multivariate} from the normalized quantities obtained from
 the Illustris-1-Dark simulation.
The distributions in this Figure are computed from $10^6$ points
generated with the multivariate gaussian. 
Similar plots for Illustris-1 and ELVIS are in the Appendix \ref{appendix:plots}.
The values for all the covariance matrices and mean vectors
foricorresponding to all the simulations are listed in the Appendix \ref{appendix:covariance}.

@@

We consider that a sampled system is similar to the M31/MW galaxy if the
distance of each of its characteristics ($w$, $c/a$, $b/a$) to the
sample mean is equal or larger than the distance of the observational
values to the sample mean.   
That is, we perform a double-tailed test using the observational
values as a threshold.

@@

>>>>

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical Dark Matter (DM) halo mass in the LG samples
and simulated baryonic effects.
However, there are three common trends.
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in the M31 galaxy have a rather typical
distribution in the LCDM context, while the MW distribution is
atypical in this respect.  
These quantitative results place the LG at the level of a $3\sigma$
outlier in the LCDM paradigm. 
We suggest that understanding the reasons for this atipicality
will require quantifying the asphericity probability distribution as a
function of halo mass and large scale environment.
The approach presented here can facilitate that kind of study and other
comparisons between different numerical setups and choices to study
satellites around LG pairs in simulations.  
\end{abstract}

@@

In the case of M31, observational studies have found a planar satellite
distribution in a special subset of $15$ satellites out of the 
total poulation of $27$ satellites \citep{2013ApJ...766..120C,
2013Natur.493...62I}, although considering the satellites ranked by
luminosity does not show any special feature.
All published studies agree on the point that the spatial distribution of
the 15 to 27 brightest satellites are consistent with spherically
symmetric distribution and easy to reproduce in LCDM simulations
\citep{2006AJ....131.1405K,2007MNRAS.374.1125M, 2013ApJ...766..120C}.
%comentario: no estoy segura de que el paper de Conn esté bien citado acá. el dice "It is clear that while the satellites of M31 when taken as a whole are no more planar than one can expect from a random distribution, a subset consisting of roughly half the sample is remarkably planar". Y en esta citación "15 to 27 brightest satellites" me parece que puede sonar contradictorio con lo que dice Conn.  
Table \ref{table:M31} and Table \ref{table:MW} summarize some more
details on the results we have just mentioned \footnote{
Those tables also include the results from this paper using the
methodolgy we describe in the upcoming sections.}.

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2006AJ....131.1405K}} & RMS width in 15 brightest
satellites & Monte Carlo satellite distributions with a power law
radial distribution & $87-99$\\
\text{\cite{2007MNRAS.374.1125M}} & $c/a$ ratio and RMS width in 16 brightest satellites & Monte Carlo from a spherical power law radial
distribution & $17$\\
\text{\cite{2013ApJ...766..120C}}& RMS width in the 27 brightest
satellites & Monte Carlo randomized satellite distribution & High\\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body Dark Matter only simulation ($\sim
10^{6}$\Msun particle mass resolution)& 56 \\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 20 halos from selected pairs in a cosmological N-body hydro simulation ($\sim
10^{6}$\Msun particle mass resolution)& 27 \\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 37 \\
\hline
\end{tabular}
\caption{Probability to find  the triaxiality and/or root mean squared
  (RMS) height of M31 satellites in LCDM simulations. 
\label{table:M31}}
\end{table*}

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2005A&A...431..517K}} & RMS width in 11 classical
satellites & Monte Carlo from a spherical power law
radial distribution. & $<0.5$ \\
\text{\cite{2005MNRAS.363..146L}} & $c/a$ ratio of 11 classical
satellites & 6 high resolution DM only N-body simulations ($\sim10^5$
\Msun\ particle mass resolution). & High\\ 
\text{\cite{2005ApJ...629..219Z}} & $c/a$ ratio and disk height in 
11 classical satellites & 3 high resolution DM only N-body
simulations. ($\sim10^6$ \Msun\ particle mass resolution) & 2 \\
\text{\cite{2007MNRAS.374.1125M}} & $c/a$ ratio and RMS width in 11-13 brightest satellites & Monte Carlo from a halo triaxiality distribution from LCDM
simulations & $<0.5$\\
\text{\cite{2009MNRAS.399..550L}}& $c/a$ ratio in 11 classical satellites & 436 halos from a
cosmological N-body simulation ($1.3\times 10^{8}$\Msun\ particle mass)
&  High \\
\text{\cite{2013MNRAS.429..725S}}& $c/a$ ratio in 12 brightest
satellites & 6 High resolution DM only N-body simulations ($\sim
10^3$\Msun particle mass resolution) of individual halos & Low \\
\text{\cite{2013MNRAS.429.1502W}}& $c/a$ ratio or RMS width in 11 brightest
satellites & 1686 halos from a cosmological DM only N-body simulation
($\sim 10^6$\Msun particle mass resolution) & 6 or 13 \\
\text{\cite{2014ApJ...789L..24P}}& $c/a$ and $b/a$ ratio in 11
classical satellites & 48 high resoltion DM only N-body simulations
($\sim 10^{5}$\Msun particle mass resolution) of both halo pairs and
isolated halos & 0.77\\
\text{\cite{2016MNRAS.457.1931S}}& $c/a$ ratio in 11 classical satellites & 12
high resolution Hydro simulation of halo pairs ($\sim 10^{4}$\Msun particle
mass resolution) & High\\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body
Dark Matter only simulation ($\sim 10^{6}$\Msun particle mass resolution)& 3\\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 20 halos from selected pairs in acosmological N-body
hydro simulation ($\sim 10^{6}$\Msun particle mass resolution)& 0.6 \\
This Work & $c/a$ ratio, $b/a$ ratio and RMS width in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 0.2 \\
\hline
\end{tabular}
\caption{Same as Table \ref{table:M31} for the MW satellites.
\label{table:MW}}
\end{table*}

@@

We control the first effect by setting as a direct point of reference
a spherical satellite distribution and not the simulations themselves.
The spherical satellite distribution is built from the data itself
(observational or simulated) by randomizing the angular position
of each satellite around the central galaxy and keeping its radial
distance fixed \citep{2017AN....338..854P}. 
We characterize the satellites in terms of the scalars describing its
deviation from the spherical distribution. 
We then build an explicit analytic probability distribution for the
asphericity; this solves the problem of having a common reference
point to compare simulations.  
We use a simulation to estimate the parameters in this distribution to
later use it as a parent sample to generate any desired number of
samples that are by construction statistically compatible with the
simulation; thus overcoming the problem of having a small number of
systems in the parent simulations.

@@

The rest of the paper describes in detail our implementation and
results. It is structured as follows. 
In Section \ref{sec:DataSamples} we list the sources of the observational and
simulated data to be used throughout the paper.
In Section \ref{sec:SpatialMeasurements} we describe the methods we
use to build halo pairs and  
characterize their satellite distributions.
In Section \ref{sec:results} we present our results to finally
conclude in Section \ref{sec:conclusions}.

@@

The base for our analysis is the catalog compiled by
\cite{2014yCat..74351928P} which reports information on all 
galaxies within 3 Mpc around the Sun to that date. 
Detailed description of the compiled catalog can be found in
\cite{2013MNRAS.435.1928P}, here we summarize the relevant features
for the current study.
The information in the catalogue is based on the catalogue compiled by
\cite{2012AJ....144....4M}.
The distance estimates are based on resolved stellar populations. 
We use three dimensional positions in a cartesian coordinate system
as computed by \cite{2013MNRAS.435.1928P}.
In this coordinate system the $z$-axis points towards the Galactic north pole, the
$x$-axis points in the direction from the Sun to the Galactic center,
and the $y$-axis points in the direction of the Galactic rotation.

For both the M31 and MW we only use the 11 to 15 brightest satellites (using
$M_V$ magnitudes) within a distance of $300$kpc from its central galaxy.
The satellites included for the MW analysis are: 
LMC, SMC, Canis Major, Sagittarius dSph, Fornax, Leo I, Sculptor,
Leo II, Sextans I, Carina, Ursa Minor, Draco, Canes Venatici (I),
Hercules and Bootes II.
The satellites included for the M31 analysis are: Triangulum, NGC205,
M32, IC10, NGC185, NGC147, Andromeda VII, Andromeda II, Andromeda
XXXII, Andromeda XXXI, Andromeda I, Andromeda VI, Andromeda XXIII, LGS 3, 
 and Andromeda III.

@@

We use publicly available data from the Illustris Project 
\citep{2014MNRAS.444.1518V}. 
This suite of cosmological simulations, performed using the quasi-Lagrangian
code AREPO \citep{2010MNRAS.401..791S}, followed the coupled evolution of dark 
matter and gas and includes parametrizations to account for the effects of
gas cooling, photoionization, star formation, stellar feedback, black
hole and super massive black hole feedback. 
The simulation volume is a cubic box with a $75$ \Mpch\ side.
The cosmological parameters correspond to a $\Lambda$CDM cosmology
consistent with WMAP-9 measurements \citep{2013ApJS..208...19H}.

@@

We buid a sample of pairs that resemble the conditions in the LG.
To construct this sample we select from Illustris-1 all the galaxies with a stellar
mass in the range $1\times10^{10}\Msun <M_{\star}<1.5 \times 10^{11}
\Msun$.
Then we select the pairs with the following conditions.

@@

In Illustris-1-Dark we use the center of mass position of the 27 pairs
in Illustris-1 to find the matching halo pairs.
After discarding the pairs with less than 15 detected subhalos in one
of the halos we end up with a total of 24 pairs in Illustris-1-Dark. 
This corresponds to a pair number density of $\sim 2 \times10^{-5}$
pairs Mpc$^{-3}$. 
Appendix A shows the physical  properties (stellar masses, maximum
circular velocities, radial velocities and separation) of those
pairs.

@@

We compute the satellite statistics for 11 up to 15 satellites.
The lower limit corresponds to the number of classical MW satellites,
while the upper limit corresponds to the minimum number of M31
satellites usually included M31 studies.
In simulations we rank the subhalos by their maximum circular
velocity, in observations we rank the satellites by its $M_V$
magnitude.

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Observations}}
    & \multicolumn{2}{c|}{\textbf{Randomized Obs.}} &
    \multicolumn{2}{c|}{\textbf{Normalized Units}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} \\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $59\pm 3$  & $21\pm 2$  & $65\pm 12$   &
    $45\pm 8$    & $-0.48\pm 0.24$ & $-2.48\pm 0.26$\\\hline
    $c/a$ ratio & $0.45\pm 0.04$ & $0.25\pm 0.05$ & $0.55\pm0.10$ &
    $0.53\pm 0.10$ & $-1.03\pm 0.37$ & $-2.18\pm 0.42$\\ \hline
    $b/a$ ratio & $0.82\pm 0.06$ & $0.80\pm 0.04$ & $0.82\pm0.07$ &
    $0.81\pm 0.08$ & $-0.02\pm 0.82$ & $-0.13\pm 0.47$\\ \hline
  \end{tabular}
  \caption{Results from observations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first column refers to the results from
    observational data, the second column uses the spherically
    randomized version of the observational data, the third column
    corresponds to the observational data recentered and normalized
    by the randomized values. \label{table:observations}}
\end{table*}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Illustris-1-Dark}} & \multicolumn{2}{c|}{\textbf{Illustris-1}} & \multicolumn{2}{c|}{\textbf{ELVIS}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW}& \textbf{M31} & \textbf{MW}\\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $62\pm 3$   & $61\pm 3$     & $70\pm 4$ & $67\pm 2$ & $70\pm 2$& $68\pm 4$ \\\hline
    $c/a$ ratio & $0.50\pm0.01$ & $0.50\pm 0.02$ & $0.52\pm 0.01$ & $0.53\pm 0.01$ & $0.54\pm 0.01$& $0.49\pm 0.02$ \\ \hline
    $b/a$ ratio & $0.79\pm0.01$ & $0.81\pm 0.02$ & $0.80\pm 0.01$ & $0.80\pm 0.02$ & $0.80\pm0.01$& $0.81\pm 0.01$\\ \hline
  \end{tabular}
  \caption{Results from simulations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first column summarizes the results from the 24 pairs in
    Illustris-1-Dark, the second column from the 20 pairs in
    Illustris-1 and the third column from the 12 pairs in the
    ELVIS project.\label{table:simulations}}
\end{table*}

@@

We compare each satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
For each quantity we compute the average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.
These normalized quantities are used to build the explicit probability
distributions for the scalars describing asphericity.

@@

After building the normalized variables with the simulated data we
perform a Kolmogorov-Smirnov test with the null hypothesis of belonging
to a normal distribution with mean and standard deviation computed
from the mean and standard deviation estimated from the data.
Although the physical quantities of interest are bound, unlike the
random variables from a normal distribution, we find that the
distributions for the normalized $w$, $c/a$ and $b/a$ are indeed
consistent with gaussian ditributions.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in observations and
  simulations (Illustris-1-Dark in this case). In all panels the
  horizontal/vertical axis corresponds to the  M31/MW or the
  most/less massive halo in the pair 
  Left. Plane width in physical units comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions (circles). 
Middle. Average from the observations (star) and the average from each
pair in the simulation (circles).
Right. Same as the middle panel except that this time each point has
been normalized (median substracted and normalized by the standard
deviation) to the results of its randomization. 
The main message of these panels is that the MW has a significantly
thinner plane both compared to the result of its own satellite
spherical randomization (left panel) and the expectation from
simulations (middle and right panel).  
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In M31 the satellite distribution is in agreement with the
expectations both from a spherical distribution and the results from the
simulations. 
\label{fig:scatter_width}}
\end{figure*}

@@

Figure \ref{fig:correlations_illustrisdm} illustrates the results from
 computing the covariance matrix and mean vector in
 Eq.\ref{eq:multivariate} from the normalized quantities obtained from
 the Illustris-1-Dark simulation.
The distributions in this Figure are computed from $10^6$ points
generated with the multivariate gaussian. 
Similar plots for Illustris-1 and ELVIS are in the Appendix \ref{appendix:plots}.
The values for all the covariance matrices and mean vectors
corresponding to all the simulations are listed in the Appendix \ref{appendix:covariance}.

@@

We consider that a sampled system is similar to the M31/MW galaxy if the
distance of each of its normalized characteristics ($w$, $c/a$, $b/a$) to the
sample mean is equal or larger than the distance of the observational
values to the sample mean.   
That is, we perform a double-tailed test using the observational
values as a threshold.

@@

Another element that influences this trend is the simulated physics.
This is evident in the comparison between Illustris-1 and Illustris-1-Dark.
These two simulations share the same characteristics except for the
inclusion of baryonic physics in Illustris-1.
We find that extreme MW systems are easier to find in the DM only
simulation. 
The results listed in Appendix \ref{appendix:covariance} show that
halos are closer to spherical once baryonic effects are included. 
For instance the mean value for the normalized width goes closer to
zero, from $-0.45\pm 0.04$  to $-0.17\pm0.04$, and the same is true
for the $c/a$ ratio that changes from $-0.61\pm0.04$ in the DM only
simulation to $-0.40\pm 0.04$ in Illustris-1.  
We postpone a detailed quantification of this effect for a future
study.
